### 2022-08-04 text/tables/tenurecount-shock-reg-fte.tex
rough first draft.

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variable: Employment Count by Tenure Group} \\ 
\cline{2-9} 
\\[-1.8ex] & \multicolumn{2}{c}{Non-tenure} & \multicolumn{2}{c}{Tenure-Track} & \multicolumn{2}{c}{Tenured} & \multicolumn{2}{c}{All} \\ 
 & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 Non-inst. Revenues & $-$0.039 & 0.590 & 0.227 & 1.108 & 0.206 & 0.227 & 0.174 & 0.454 \\ 
  & (0.084) & (0.826) & (0.162) & (0.829) & (0.096) & (0.200) & (0.095) & (0.265) \\ 
  Tuition Revenue & 0.069 & $-$0.166 & 0.088 & $-$0.229 & 0.071 & 0.063 & 0.027 & $-$0.074 \\ 
  & (0.109) & (0.370) & (0.109) & (0.355) & (0.038) & (0.106) & (0.054) & (0.158) \\ 
 \hline \\[-1.8ex] 
Observations & 4,825 & 4,825 & 5,094 & 5,094 & 5,130 & 5,130 & 5,181 & 5,181 \\ 
R$^{2}$ & 0.849 & 0.843 & 0.780 & 0.743 & 0.902 & 0.902 & 0.935 & 0.928 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}