### 2022-08-09 text/tables/tenurecount-shock-reg-fte.tex
extra detail.

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variable: Employment Count by Tenure Group} \\ 
\cline{2-9} 
\\[-1.8ex] & \multicolumn{2}{c}{Non-tenure} & \multicolumn{2}{c}{Tenure-Track} & \multicolumn{2}{c}{Tenured} & \multicolumn{2}{c}{All} \\ 
 & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 State Funding & $-$0.0004 & 1.499 & 0.048 & 1.926 & 0.051 & 0.267 & 0.037 & 0.893 \\ 
  & (0.016) & (10.764) & (0.051) & (8.050) & (0.029) & (1.104) & (0.023) & (4.610) \\ 
 \hline \\[-1.8ex] 
Observations & 4,825 & 4,825 & 5,094 & 5,094 & 5,130 & 5,130 & 5,181 & 5,181 \\ 
R$^{2}$ & 0.849 & 0.588 & 0.777 & $-$0.413 & 0.898 & 0.872 & 0.932 & 0.426 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}