### 2022-11-20 text/sections/09-appendix.tex
Updated draft.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Appendix section
% Set-up the section.
\newpage
\appendix
\setcounter{table}{0}
\renewcommand{\thetable}{A\arabic{table}}
\setcounter{figure}{0}
\renewcommand{\thefigure}{A\arabic{figure}}

% Start appendix
\section{Appendix}
\label{appendix}
This project used data which are fully public, and computational tools which are fully open-source.
As such, all code and data involved in this project are available at this project's Github repository, available at \url{https://github.com/shoganhennessy/state-faculty-composition}.
They may be used for replication, or as the basis for further work, as needed.
Any comments or suggestions may be sent to me at \href{mailto:seh325@cornell.edu}{\nolinkurl{seh325@cornell.edu}}, or raised as an issue on the Github project.

A number of statistical packages, for the $R$ language \citep{R2022}, made the empirical analysis for this paper possible.
\begin{itemize}
    \item \textit{Tidyverse} \citep{tidyverse} collected tools for data analysis in the R language.
    \item \textit{LFE} \citep{lfe} implemented linear fixed effect models, with instruments, crucial for the empirical estimation in \autoref{sec:empirics}.
    \item \textit{Stargazer} \citep{stargazer} provided a method to efficiently convert empirical results into presentable output in \LaTeX.
    \item \textit{Lpirfs} \citep{lpirfs2019} implemented estimation of the \cite{jorda2005} local projections methods, with instrumental variables, crucial to the local projections estimates presented in this project.
\end{itemize}

\subsection{First Stage Estimates, Individual Outcomes}
\label{appendix:part1}

\begin{table}[h!]
    \singlespacing
    \centering
    \caption{IBHED Summary Statistics, Entire Professor Panel 2010--2021.}
    \makebox[\textwidth][c]{\input{tables/illinois-summary-rolling.tex}}
    \label{tab:illinois-summary-rolling}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: This table presents the summary statistics for professors with an observed year of first-year at their university.
    \end{flushleft}
\end{table}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{First Stage Estimates, for Total University Revenues at the Individual-Level.}
    \makebox[\textwidth][c]{\input{tables/firststage-illinois-rolling.tex}}
    \label{tab:firststage-illinois}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the university-year level.
    \end{flushleft}
\end{table}

\begin{figure}[h!]
    \centering
    \singlespacing
    \caption{Local Projection Estimates for First-Stage \autoref{eqn:secondstage1_indiv}.}
    \includegraphics[width=0.6\textwidth]{figures/firststage-illinois-lp-rolling.png}
    \label{fig:firststage-illinois-lp}
\end{figure}

\subsection{Additional Results, University-Level}

\autoref{tab:facultysalaries-shock-reg} presents OLS and IV estimates where the outcome is (log) mean salary for professors employed at the university, separated by position.
Similarly, \autoref{fig:all-salaries-lp} shows the staying-power via projection methods.
The outcome, mean salary is crude.

\begin{table}[H]
    \singlespacing
    \centering
    \caption{OLS and 2SLS Estimates for University Faculty-Tenure Composition.}
    \makebox[\textwidth][c]{\input{tables/tenurecount-shock-reg-fte.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the university-year level.
    \end{flushleft}
    \label{tab:tenurecount-shock-reg-fte}
\end{table}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{OLS and 2SLS Estimates for University Faculty Salaries.}
    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-reg-fte.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the state-year level.
    \end{flushleft}
    \label{tab:facultysalaries-shock-reg}
\end{table}

\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Local Projection Estimates for Professor Salary, Mean within Each University.}
    \includegraphics[width=0.6\textwidth]{figures/all-salaries-lp.png}
    \label{fig:all-salaries-lp}
\end{figure}

\subsection{Additional Results, Individual-Professor Level}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Salaries at Illinois Universities.}
    \makebox[\textwidth][c]{\input{tables/facultysalaries-shock-illinois.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution-year level.
    \end{flushleft}
    \label{tab:facultysalaries-shock-illinois}
\end{table}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities, using Rolling Instrument.}
    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois-rolling.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
    \end{flushleft}
    \label{tab:promotion-shock-illinois-rolling}
\end{table}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Promotion Rate at Illinois Universities, using Base-Year Instrument.}
    \makebox[\textwidth][c]{\input{tables/promotion-shock-illinois.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
    \end{flushleft}
    \label{tab:promotion-shock-illinois}
\end{table}

\begin{figure}[H]
    \centering
    \singlespacing
    \caption{Local Projection Estimates for Promotion Rate at Illinois Universities.}
    \includegraphics[width=0.6\textwidth]{figures/promoted-illinois-lp-rolling.png}
    \label{fig:promoted-illinois-lp}
\end{figure}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities, using Rolling Instrument.}
    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois-rolling.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
    \end{flushleft}
    \label{tab:facultyleaving-shock-illinois-rolling}
\end{table}

\begin{table}[H]
    \singlespacing
    \centering
    \caption{2SLS Estimates for Faculty Exit Rate at Illinois Universities, using Base-Year Instrument.}
    \makebox[\textwidth][c]{\input{tables/facultyleaving-shock-illinois.tex}}
    \begin{flushleft}
        \footnotesize
        \textbf{Note}: Standard errors are clustered at the institution and first year of employment level.
    \end{flushleft}
    \label{tab:facultyleaving-shock-illinois}
\end{table}