### 2023-07-11 text/tables/tenurecount-shock-reg-fte.tex

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variable: Employment Count by Tenure Group} \\ 
\cline{2-9} 
\\[-1.8ex] & \multicolumn{2}{c}{Non-tenure} & \multicolumn{2}{c}{Tenure-Track} & \multicolumn{2}{c}{Tenured} & \multicolumn{2}{c}{All} \\ 
 & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 State Funding & 0.003 & 3.332 & 0.045 & 2.367 & 0.051 & 0.886 & 0.043 & 2.925 \\ 
  & (0.017) & (66.670) & (0.049) & (15.854) & (0.028) & (6.337) & (0.029) & (43.349) \\ 
 \hline \\[-1.8ex] 
Observations & 4,633 & 4,633 & 4,851 & 4,851 & 4,878 & 4,878 & 4,906 & 4,906 \\ 
R$^{2}$ & 0.849 & $-$0.419 & 0.782 & $-$1.185 & 0.896 & 0.470 & 0.927 & $-$5.366 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}