### 2023-07-18 text/tables/facultyhires-illinois-reg-rawcount.tex

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variable: Employment Count by Professor Group} \\ 
\cline{2-9} 
\\[-1.8ex] & \multicolumn{2}{c}{Lecturer} & \multicolumn{2}{c}{Assistant} & \multicolumn{2}{c}{Full} & \multicolumn{2}{c}{All} \\ 
 & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 State Funding & $-$0.526 & 1.383 & 0.252 & 0.529 & $-$0.065 & 0.058 & $-$0.977 & 0.491 \\ 
  & (0.712) & (0.695) & (0.597) & (0.670) & (0.354) & (0.377) & (1.345) & (1.869) \\ 
 \hline \\[-1.8ex] 
Outcome Mean & 73.275 & 73.275 & 42.771 & 42.771 & 12.301 & 12.301 & 151.932 & 151.932 \\ 
Observations & 131 & 131 & 131 & 131 & 113 & 113 & 132 & 132 \\ 
R$^{2}$ & 0.805 & 0.795 & 0.921 & 0.921 & 0.762 & 0.762 & 0.907 & 0.906 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}