### 2023-07-18 text/tables/firststage-balance-count.tex

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variables: University Characteristics} \\ 
\cline{2-9}
\\[-1.8ex] & Enrolment & State & Total & Non-inst. & Lecturers & Assistant & Full & All \\ 
\\[-1.8ex] &           & Funding & Revenues & Revenues & Lecturers & Professors & Professors & Professors \\
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 Funding & 0.517   & $-$9,217 & $-$7,633 & 886 & 0.007 & $-$0.001 & $-$0.003 & 0.004 \\ 
 Shock   & (0.199) & (2,452) & (12,771) & (4,093) & (0.002) & (0.001) & (0.003) & (0.005) \\ 
 \hline \\[-1.8ex] 
Uni. + Year \\
fixed effects? & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\
Observations & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 \\ 
R$^{2}$ & 0.960 & 0.950 & 0.881 & 0.932 & 0.745 & 0.886 & 0.973 & 0.954 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}