### 2023-12-19 text/tables/firststage-balance-count.tex

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variables: University Characteristics} \\ 
\cline{2-9} 
\\[-1.8ex] & Enrolment & State Funding & Total revenues & Non-inst. revenues & Lecturers & Assistant professors & Full professors & All professors \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 State Funding & 0.517 & $-$9,217.091 & $-$7,633.377 & 886.006 & 0.007 & $-$0.001 & $-$0.003 & 0.004 \\ 
  & (0.199) & (2,451.719) & (12,771.130) & (4,092.780) & (0.002) & (0.001) & (0.003) & (0.005) \\ 
 \hline \\[-1.8ex] 
Uni. + Year fixed effects? & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 \\ 
R$^{2}$ & 0.960 & 0.950 & 0.881 & 0.932 & 0.745 & 0.886 & 0.973 & 0.954 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}