### 2023-12-19 text/tables/firststage-balance-log.tex

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variables: University Characteristics} \\ 
\cline{2-9} 
\\[-1.8ex] & Enrolment & State Funding & Total revenues & Non-inst. revenues & Lecturers & Assistant professors & Full professors & All professors \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 State Funding & 0.107 & $-$0.870 & 0.031 & $-$0.091 & 0.534 & $-$0.024 & $-$0.027 & 0.043 \\ 
  & (0.037) & (0.079) & (0.082) & (0.063) & (0.122) & (0.066) & (0.034) & (0.038) \\ 
 \hline \\[-1.8ex] 
Uni. + Year fixed effects? & Yes & Yes & Yes & Yes & Yes & Yes & Yes & Yes \\ 
Observations & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 & 17,012 \\ 
R$^{2}$ & 0.976 & 0.914 & 0.979 & 0.975 & 0.782 & 0.902 & 0.963 & 0.965 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}