### 2023-12-19 text/sections/03-concepts.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Section on possible outcomes
\section{Conceptual Framework}
\label{sec:conceptual}

Universities can respond multiple ways to the falls in per student state funding.
\cite{NBERw23736} established that public universities who experienced state funding cuts did not fully off-set by raising tuition, and cutting total spending.
But it is not clear how these spending cuts went on to affect faculty --- or whether they affected faculty at all.

Wider changes in the American higher education sector may concurrently explain the trends in state funding and faculty substitution.
College selectivity changed drastically over the 21st century, where the top universities have become more selective while the average university less selective \citep{hoxby2009changing}.
While the average public university is becoming less selective, it is possible that their most productive or research-focused faculty more often move to more selective and prestigious (and often private) institutions, leaving public universities substituting toward lecturers over time.
This is one way in which the trends may be concurrent, but not causal.
I address this issue by using shift-share instrument approach to isolate funding cuts that affected more funding-intensive universities, measuring the rate of substitution between lecturers and professors in response to yearly changes in state-wide funding.

Universities could respond to funding cuts by cutting faculty salaries, leading to more faculty leaving for jobs at other universities.
Multiple universities passed a university-wide pay-cut for their faculty in response to state budget cuts around the 2008 recession, and Cornell University implemented a professor salary cut in 2020.\footnote{
    The Cornell University administration expected large fiscal squeezes in mid-2020, so imposed a faculty salary cut, and then returned the amount cut later in 2020 when the financial squeeze did not materialise.}
Funding cuts could also limit the amount public universities can offer to new faculty hires, leading to fewer or less accomplished faculty accepting these offers.\footnote{
    Faculty often consider their outside options in job decisions, considering offers for employment and/or promotion from multiple universities at the same time (see \citealt{blackaby2005} for an overview of faculty outside options).
    If public universities can only make low salary offers thanks to funding cuts, then they are less likely to land professors with multiple offers.
}
This effect may not be the same across faculty seniority; most universities hire more junior faculty each year, so that yearly hiring disruptions may disrupt the looser market for established full professors more than tighter market for assistant professors or lecturers.
While these factors make it unclear who will be affected most by funding cuts, there is one empirical fact worth noting: the average Illinois assistant professor is paid $\$76,897$, tenured professor $\$109,283$, while lecturers earn $\$31,449$ more than double that of a lecturer (annual salary for 2010--2021 in 2021 USD, see \autoref{tab:illinois-summary}).
Secondly, lecturers are (by definition) granted shorter contracts than professors; assistant professors mostly work on 5--6 year long contracts, full professors on indefinite contracts (tenure), lecturers on yearly or term-wise contracts.
If a public university's primary obligation is to teach, and they must fulfil this objective with fewer resources, substituting away from tenure-track and tenured professors towards lecturers is rationalisable.

It is not clear when the effects of funding cuts will be realised.
University departments often employ faculty on multiple-year contracts (e.g., the tenure contract), and individual professors' decisions to retire, move university, or leave academia are multiple-year commitments, so that it may take multiple years for university funding to change the faculty composition, or trickle down to faculty.
For example, university departments coordinate hiring in annual cycles, so that a budget short-fall in year $t$ may have no effect on the in-progress hiring committee and resulting hiring decisions, yet the hiring cycle in year $t+1$ may be postponed or cancelled.
As such, this paper investigates the dynamics of state funding shocks' effects on faculty, for the years following a state funding cut.