### 2024-05-08 text/tables/facultycount-illinois-reg-rawcount.tex

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variable: Employment Count by Professor Group} \\ 
\cline{2-9} 
\\[-1.8ex] & \multicolumn{2}{c}{Lecturers} & \multicolumn{2}{c}{Asst. Professors} & \multicolumn{2}{c}{Full Professors} & \multicolumn{2}{c}{All Faculty} \\ 
 & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 State Funding & $-$3.729 & 0.376 & 0.415 & 3.208 & $-$1.269 & 0.117 & $-$4.084 & 3.890 \\ 
  & (1.721) & (1.725) & (1.728) & (2.493) & (1.548) & (1.200) & (4.352) & (5.668) \\ 
 \hline \\[-1.8ex] 
Outcome Mean & 351.306 & 351.306 & 273.757 & 273.757 & 477.597 & 477.597 & 1303.014 & 1303.014 \\ 
Observations & 144 & 144 & 144 & 144 & 144 & 144 & 144 & 144 \\ 
R$^{2}$ & 0.886 & 0.882 & 0.970 & 0.969 & 0.991 & 0.991 & 0.981 & 0.980 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}