### 2024-05-08 text/tables/facultyhires-illinois-reg-rawcount.tex

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variable: Yearly New Hires by Professor Group} \\ 
\cline{2-9} 
\\[-1.8ex] & \multicolumn{2}{c}{Lecturers} & \multicolumn{2}{c}{Asst. Professors} & \multicolumn{2}{c}{Full Professors} & \multicolumn{2}{c}{All Faculty} \\ 
 & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 State Funding & $-$2.551 & $-$2.280 & 0.060 & $-$5.126 & $-$0.095 & $-$2.562 & $-$2.677 & 23.012 \\ 
  & (1.420) & (9.168) & (0.675) & (7.210) & (0.249) & (7.051) & (2.185) & (52.652) \\ 
 \hline \\[-1.8ex] 
Outcome Mean & 73.275 & 73.275 & 42.771 & 42.771 & 12.301 & 12.301 & 151.932 & 151.932 \\ 
Observations & 131 & 131 & 131 & 131 & 113 & 113 & 132 & 132 \\ 
R$^{2}$ & 0.839 & 0.839 & 0.934 & 0.902 & 0.788 & 0.752 & 0.918 & 0.793 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}