### 2024-05-08 text/tables/parttime-shock-reg-fte.tex

\begin{tabular}{@{\extracolsep{5pt}}lcccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{6}{c}{Dependent Variable: Log Lecturer Count per Students, by Position} \\ 
\cline{2-7} 
\\[-1.8ex] & \multicolumn{2}{c}{Part-time Lecturers} & \multicolumn{2}{c}{Full-time Lecturers} & \multicolumn{2}{c}{All Lecturers} \\ 
 & OLS & 2SLS & OLS & 2SLS &  &  \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6)\\ 
\hline \\[-1.8ex] 
 State Funding & $-$0.042 & $-$0.258 & $-$0.045 & $-$0.217 & 0.023 & 0.055 \\ 
  & (0.133) & (0.369) & (0.129) & (0.317) & (0.103) & (0.104) \\ 
 \hline \\[-1.8ex] 
Outcome Mean & 2.052 & 2.052 & 1.025 & 1.025 & 0.654 & 0.654 \\ 
Observations & 6,077 & 6,077 & 6,077 & 6,077 & 6,077 & 6,077 \\ 
R$^{2}$ & 0.712 & 0.711 & 0.672 & 0.671 & 0.788 & 0.788 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}