### 2024-09-26 text/tables/facultycount-replicated-corrected-1990s.tex
Added a number of new things.  2 minor comments to address before sub.

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variable: Log Faculty Count per Students, by Position} \\ 
\cline{2-9} 
\\[-1.8ex] & \multicolumn{2}{c}{Lecturers} & \multicolumn{2}{c}{Asst. Professors} & \multicolumn{2}{c}{Full Professors} & \multicolumn{2}{c}{All Faculty} \\ 
 & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 State Funding & $-$0.062 & $-$0.739 & 0.075 & 0.142 & 0.086 & 0.135 & 0.060 & 0.051 \\ 
  & (0.031) & (0.149) & (0.042) & (0.073) & (0.045) & (0.099) & (0.039) & (0.052) \\ 
 \hline \\[-1.8ex] 
Outcome Mean & 0.526 & 0.526 & 1.255 & 1.255 & 2.409 & 2.409 & 4.245 & 4.245 \\ 
Observations & 10,285 & 10,285 & 10,285 & 10,285 & 10,285 & 10,285 & 10,285 & 10,285 \\ 
R$^{2}$ & 0.734 & 0.662 & 0.739 & 0.737 & 0.807 & 0.805 & 0.818 & 0.818 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}