### 2024-09-26 text/tables/facultycount-replicated-corrected-2010s.tex

\begin{tabular}{@{\extracolsep{5pt}}lcccccccc} 
\\[-1.8ex]\hline 
\hline \\[-1.8ex] 
 & \multicolumn{8}{c}{Dependent Variable: Log Faculty Count per Students, by Position} \\ 
\cline{2-9} 
\\[-1.8ex] & \multicolumn{2}{c}{Lecturers} & \multicolumn{2}{c}{Asst. Professors} & \multicolumn{2}{c}{Full Professors} & \multicolumn{2}{c}{All Faculty} \\ 
 & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS & OLS & 2SLS \\ 
\\[-1.8ex] & (1) & (2) & (3) & (4) & (5) & (6) & (7) & (8)\\ 
\hline \\[-1.8ex] 
 State Funding & 0.224 & 0.177 & 0.209 & 0.174 & 0.228 & 0.098 & 0.211 & 0.128 \\ 
  & (0.092) & (0.115) & (0.045) & (0.073) & (0.063) & (0.052) & (0.050) & (0.052) \\ 
 \hline \\[-1.8ex] 
Outcome Mean & 0.644 & 0.644 & 1.101 & 1.101 & 2.128 & 2.128 & 3.949 & 3.949 \\ 
Observations & 6,727 & 6,727 & 6,727 & 6,727 & 6,727 & 6,727 & 6,727 & 6,727 \\ 
R$^{2}$ & 0.828 & 0.828 & 0.827 & 0.827 & 0.910 & 0.908 & 0.931 & 0.929 \\ 
\hline 
\hline \\[-1.8ex] 
\end{tabular}